Crunching Large Graphs with Commodity Processors

### 2011-01-26 doc/hotpar2011/softxmt-hotpar2011.tex
modified:   softxmt-hotpar2011.tex

\authorinfo{Jacob Nelson, Brandon Myers, A.H. Hunter, Preston Briggs,
  Dan Grossman, Mark Oskin, Carl Ebeling, Simon Kahan, Luis
  Ceze}{University of Washington}{\{nelson, bdmyers, ahh, preston,
  djg, oskin, ebeling, skahan, luisceze\}@cs.washington.edu}

@@

% Whereas in the PGAS model coherency of global copies is implicitly
% maintained across the system, % \todo{cite pgas?}
% we completely forego
% caching of globally shared data.

@@

%We minimize context size by treating context switches as function
%calls, as in~\cite{charm}. This allows us 
%to save and restore the
%minimum number of registers allowed by the ABI, and 
%Simon:  not really true -- we could save fewer if we knew which of the
%callee save registers weren't going to be clobbered.
We treat context switches as function calls, as in~\cite{charm}. This allows us 
to depend on the compiler to save and restore caller-save registers while we explicitly save and restore all the callee-save registers.  We minimize
context switch time by doing all switching and scheduling in user space.

@@

%\paragraph{Multi-node ideas.}
In a multi-node implementation, synchronization operations can be delegated to a memory manager
thread or accelerator. Synchronization on remote data would be
performed by the remote memory manager. This may simplify the
implementation: if only a single core (or single accelerator) is
accessing the data being synchronized, the use of memory fences may be
reduced or eliminated.

@@

Our evaluation consists of two parts. First, we demonstrate that the
runtime system can achieve the same performance as when 
explicit memory concurrency is available. Second, we look at the runtime
within the multi-node picture and show that it can tolerate large
latencies. In each part we begin by studying relevant aspects of our
test machine's memory system so we can interpret our runtime results.

@@

Our goal of presenting a global view of distributed memory to the
programmer is shared by the PGAS community, and is used in languages
like Chapel \cite{chapel}, X10 \cite{X10}, and UPC \cite{upc}. They
obtain performance by minimizing references to remote nodes, whereas we
design for remote references.

>>>>

\authorinfo{Jacob Nelson$^{\dagger}$, Brandon Myers$^{\dagger}$, A.H. Hunter$^{\dagger}$, Preston Briggs$^{\dagger}$,
  Dan Grossman$^{\dagger}$, Mark Oskin$^{\dagger}$, Carl Ebeling$^{\dagger}$, Simon Kahan$^{{\dagger \ddagger}}$, Luis
  Ceze$^{\dagger}$}{\textdagger University of Washington, \textdaggerdbl Pacific Northwest National Laboratory}{\{nelson, bdmyers, ahh, preston,
  djg, oskin, ebeling, skahan, luisceze\}@cs.washington.edu}

@@

% Whereas in the PGAS model coherency of global copies is implictly
% maintained across the system, % \todo{cite pgas?}
% we completely forego
% caching of globally shared data.

@@

%We minimize context size by treating context switches as function
%calls, as in~\cite{charm}. This allows us 
%to save and restore the
%minimum number of registers allowed by the ABI, and 
%Simon:  not really true -- we could save fewer if we knew which of the
%callee save registers weren't going to be clobbered.
We treat context switches as function calls, as in~\cite{charm}. This allows us 
to depend on the compiler to save and restore caller-save registers while we explicitly save and restore all the callee-save registers:  the size of a coroutine context is hardly more than the stack space needed to save these registers.  We minimize
context switch time by doing all switching and scheduling in user space (kernel traps are at least an order of magnitude slower.)

@@

%\paragraph{Multi-node ideas.}
In a multi-node implementation, synchronization operations can be delegated to a memory manager
thread or accelerator. Sychronization on remote data would be
performed by the remote memory manager. This may simplify the
implementation: if only a single core (or single accelerator) is
accessing the data being synchronized, the use of memory fences may be
reduced or eliminated.

@@

Our evaulation consists of two parts. First, we demonstrate that the
runtime system can achieve the same performance as when 
explicit memory concurrency is available. Second, we look at the runtime
within the multi-node picture and show that it can tolerate large
latencies. In each part we begin by studying relevant aspects of our
test machine's memory system so we can interpret our runtime results.

@@

Our goal of presenting a global view of distributed memory to the
programmer is shared by the PGAS community, and is used in languages
like Chapel \cite{chapel}, X10 \cite{X10}, and UPC \cite{upc}. They
obtain performance by minimizing references to remote nodes, whereas we
optimize for this case.
